---
title: "Global Terrorism Prediction"
author: "Yuan Xiang, Jingyu Gao, Lipei Tao"
date: "5/8/2017"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective

Analyses of terrorist groups and incidents have begun to recognize the evolutionary nature of terrorism, as well as its ability to adapt to counter-terrorism tactics. This project is intended to predict the clusters of terrorist attack using attack type, weapon type, target type, suicide dummy, month. The analysis is performed on 4 countries that have the most frequent attacks: Iraq, Afghanistan, Pakistan, India.


## Data Used
Our data is the Global Terrorism Database compiled by the University of Maryland. It collects information about each terrorist attack reported by governments around the world since 1970. The latest update has over 15,0000 records. Key variables of this dataset include:
1.	Types of attack: bombing, kidnapping, etc
2.	Weapon used in the attack: Handgun, Automatic Weapon, TNT, etc.
3.	Time: year, month, date
4.	Location: country/region, state/province, city; geographical coordinates
5.	Identification of attackers: recognized terrorist group, individuals
6.	Types of weapon used in the attack
7.	Types of targets: civilian, government agencies, military bases, etc


## Data Preparation and Cleaning
```{r comment=NA, cache=FALSE, message=FALSE}
# load data
load("/Users/yuanxiang/Documents/Project/global-terrorism/data/globalterror.RData")
project_data <- globalterror

new <- c("eventid","country_txt", "latitude","longitude","attacktype1", "targtype1", 
         "region", "weaptype1","imonth",
          "suicide","gname","guncertain1","doubtterr" )

tocluster.data <- project_data[new]
head(tocluster.data)
```

```{r comment=NA, cache=FALSE, message=FALSE}
# clean data
tocluster.data <- tocluster.data[tocluster.data$guncertain1 == 0 | tocluster.data$doubtterr == 0,] 
tocluster.data <- tocluster.data[tocluster.data$gname != "Unknown",]
tocluster.data <- tocluster.data[!tocluster.data$weaptype1 %in% c(12, 13),]
tocluster.data <- tocluster.data[tocluster.data$attacktype1 != 9, ]
tocluster.data <- tocluster.data[!tocluster.data$targtype1 %in% c(13,20), ]

# Convert data into categorical variables
tocluster.data[, 5:ncol(tocluster.data)] <- lapply(tocluster.data[, 5:ncol(tocluster.data)], as.factor)

str(tocluster.data)
```


## Hierarchical Cluster

### Write Functions to perform hierarchical clusters
```{r comment=NA, cache=FALSE, message=FALSE}
#HC Object function
hc.object <- function(country, featurelist){
  # Args:
      #       country: country dataframe
      #       featurelist: intput features vectors
      #
  # Returns:
      #       A HC object 
  
  library(cluster)
  country.gower <- daisy(country[featurelist], metric = "gower")
  country.gmat <- as.matrix(country.gower)
  hc <- hclust(country.gower, method="ward.D")
  return(hc)
}

#Group function
set.groups <- function(hc.object, klusters){
  # Args:
      #       hc.object: HC object
      #       kclusters: the numbers of clusters
      #
  # Returns:
      #       Groups for the passed HC object 
  library(cluster)
  groups <- cutree(hc.object, k = klusters)
  return(groups)
}

#Maps function
draw.map <- function(country, groups){
  # Args:
      #       country: country dataframe
      #       groups: clusters given by hclust function
      #
  # Returns:
      #       Map for the countries passed in
  library(maps)
  country.map <- map("world", region = country, interior=FALSE)
  country.map <- map("world",boundary=FALSE,col="black",add=TRUE)
  country.map <- points(eval(parse(text=country))$longitude, eval(parse(text=country))$latitude, pch = 19, cex=0.3,col=factor(groups))
  legend(x="topleft", bty = "n", legend=levels(factor(groups)), 
         cex = 1,  x.intersp=0, xjust=0, yjust=0, text.col=seq_along(levels(factor(groups))))
  return(country.map)
}
```


### Prepare data for hierarchical clustering methods
```{r comment=NA, cache=FALSE, message=FALSE}
# Create country data frames
Iraq <- tocluster.data[tocluster.data$country_txt=="Iraq",]
Pakistan <- tocluster.data[tocluster.data$country_txt=="Pakistan",]
Afghanistan <- tocluster.data[tocluster.data$country_txt=="Afghanistan",]
India <- tocluster.data[tocluster.data$country_txt=="India",]
feature.list <- c("attacktype1", "targtype1", "weaptype1","imonth", "suicide")

```


### Apply hierarchical clustering methods
```{r comment=NA, cache=FALSE, message=FALSE}
#Loop through countries and numbers of k
country.list <- c("Iraq", "Pakistan", "Afghanistan", "India")
k.list <- c(3:6)

for(i in country.list){
  df.hc <- hc.object(eval(parse(text=i)), feature.list)
  print(paste(i))
  assign(paste0(i,".hc"), df.hc)
  df.plot <- plot(eval(parse(text = paste0(i,".hc"))))
  dev.print(pdf, paste0(i,".hc",".pdf"))
  assign(paste0(i,".plot"), df.plot)
  carto.data <- cbind(eval(parse(text=i))$longitude, eval(parse(text=i))$latitude)
  for(j in k.list){
    df.k <- set.groups(eval(parse(text = paste0(i,".hc"))), j)
    assign(paste0(i,".k", j), df.k)
    draw.map(i, eval(parse(text = paste0(i,".k", j))))
    dev.print(pdf, paste0(i,".map", ".k",j,".pdf"))
    carto.data <- cbind(carto.data, eval(parse(text = paste0(i,".k", j))))
  }
  assign(paste0(i, ".carto"), carto.data)
}
```

```{r comment=NA, cache=FALSE, message=FALSE}
# rename the output matrix
colnames(Afghanistan.carto) <- c("longitude","latitude","k3","k4","k5","k6")
colnames(India.carto) <- c("longitude","latitude","k3","k4","k5","k6")
colnames(Iraq.carto) <- c("longitude","latitude","k3","k4","k5","k6")
colnames(Pakistan.carto) <- c("longitude","latitude","k3","k4","k5","k6")
colnames(carto.data) <- c("longitude","latitude","k3","k4","k5","k6")
```


### Prepare dataframe for the summary table of clusters
```{r comment=NA, cache=FALSE, message=FALSE}
# Convert output to dataframe and rename the output
Afghanistan.carto <- as.data.frame(Afghanistan.carto)
India.carto <- as.data.frame(India.carto)
Pakistan.carto <- as.data.frame(Pakistan.carto)
Iraq.carto <- as.data.frame(Iraq.carto)
carto.data <- as.data.frame(carto.data)

# rename the columns
colnames(Afghanistan.carto) <- c("longitude","latitude","k3","k4","k5","k6")
colnames(India.carto) <- c("longitude","latitude","k3","k4","k5","k6")
colnames(Iraq.carto) <- c("longitude","latitude","k3","k4","k5","k6")
colnames(Pakistan.carto) <- c("longitude","latitude","k3","k4","k5","k6")
colnames(carto.data) <- c("longitude","latitude","k3","k4","k5","k6")

# append groups to the country dataframe(k = 3)
Afghanistan.group <- cbind(Afghanistan[5:9], Afghanistan.carto$k3)
India.group <- cbind(India[5:9], India.carto$k3)
Pakistan.group <- cbind(Pakistan[5:9], Pakistan.carto$k3)
Iraq.group <- cbind(Iraq[5:9], Iraq.carto$k3)

# rename cluster column
colnames(Afghanistan.group)[ncol(Afghanistan.group)] <- c("group")
colnames(India.group)[ncol(India.group)] <- c("group")
colnames(Pakistan.group)[ncol(Pakistan.group)] <- c("group")
colnames(Iraq.group)[ncol(Iraq.group)] <- c("group")

```


### What do each clusters represent
```{r comment=NA, cache=FALSE, message=FALSE}
# Summary of clusters
# Create mode function
Mode <- function(x){
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# Get mode features for each clusters
overview.Iraq <- aggregate(Iraq.group, list(Iraq.group$group), Mode)
overview.India <- aggregate(India.group, list(India.group$group), Mode)
overview.Pakistan <- aggregate(Pakistan.group, list(Pakistan.group$group), Mode)
overview.Afghanistan <- aggregate(Afghanistan.group, list(Afghanistan.group$group), Mode)
```

Take a look at the result
```{r comment=NA, cache=FALSE, message=FALSE}
overview.Afghanistan
overview.India
overview.Iraq
overview.Pakistan
```


## Choose Clusters
We looked at the dendrogram to decide the possible values of k, and loop through the possible choices of k, and plot it into the maps to find the best number of clusters. We found that k = 3 yields the most significant patterns.


## Results
### Afghanistan 
* Geographically
    + The attack is generally randomly distributed

* Meaning of the clusters
    + In Cluster 1, most of the attack type is bombing, the main target     is police, the main weapon type is explosive bomb, the most frequent     month is May, it doesn't use suicide attack.

    + In cluster 2, most of the attack type is armed assault, the main      target is police, the main weapon type is firearms, the most   
    frequent month is May, it doesn't use suicide attack.
    
    + In cluster 3, most of the attack type is armed assault, the main      target is government, the main weapon type is firearms, the most   
    frequent month is August, it doesn't use suicide attack.
    

### Pakistan
* Geographically
    + Cluster 1 is located in the south of the country
    + Cluster 2 is concentrated on the northeast of the country

* Meaning of the clusters
    + In Cluster 1, most of the attack type is bombing, the main target     is military, the main weapon type is explosive bomb, the most 
    frequent month is February, it doesn't use suicide attack

    + In cluster 2, most of the attack type is armed assault, the main      target is military, the main weapon type is firearms, the most   
    frequent month is May, it doesn't use suicide attack
    
     + In cluster 3, most of the attack type is bombing, the main      
     target is police, the main weapon type is explosive bomb, the most 
     frequent month is February, it uses suicide attack

    
### India
* Geographically
    + The overall attacks are concentrated on the northeast of the 
    India.
    + Cluster 3 is mostly concentrated in the Odisha and Chattisgarh
    + Cluster 1 is located in Jharkhand and the central Manipur

* Meaning of the clusters
    + In Cluster 1, most of the attack type is bombing, the main target     is police, the main weapon type is explosive bomb, the most frequent     month is April, it doesn't use suicide attack

    + In cluster 2, most of the attack type is facility/infrastructure      attack, the main target is private business, the main weapon type is     incendiary, the most frequent month is June, it doesn't use suicide 
    attack
    
    + In cluster 3, most of the attack type is armed assault, the main 
    target is private citizens, the main weapon type is firearms, the 
    most frequent month is April, it doesn't use suicide attack


### Iraq
* Geographically
    + Most attacks belong to Cluster 2
    + Most attacks are along the road 
    + Most Cluster 3 are distributed along the road to Mosul

* Meaning of the clusters
    + In Cluster 1, most of the attack type is bombing, the main target     is police, the main weapon type is explosive bomb, the most frequent     month is December, it uses suicide attacks

    + In cluster 2, most of the attack type is bombing, the main target 
    is private citizens, the main weapon type is explosive bomb, the 
    most frequent month is July, it doesn't use suicide attack.
    
    + In cluster 3, most of the attack type is armed assault, the main 
    target is military, the main weapon type is firearms, the 
    most frequent month is June, it doesn't use suicide attack





